---
title: "Getting started with directsrr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with directsrr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment="#>")
library(directsrr)
```

## Overview

`directsrr` computes directly standardised rates (DSR) with confidence intervals (CI) and compares groups using standardised rate ratios (SRR) and standardised rate differences (SRD).

The main function, `directsrr()`, implements the gamma-based interval of Tiwari, Clegg, and Zou (2006) and the Dobson et al. (1991) interval. The package is adapted from the Stata command `distrate` (Consonni et al. 2012), but uses an ordinary data frame with the standard population supplied as a column. It also supports standardisation by multiple variables such as age and sex, and grouping by one or more variables.

---

## Case 1 — Standardise by age, compare one grouping variable

The simplest and most common use: standardise by a single age column and
compare two or more groups.

```{r}
data(directsrr_ExampleData)
directsrr_ExampleData
```

```{r}
directsrr(
  data        = directsrr_ExampleData,
  event       = "events",
  pop         = "population",
  stdpop      = "stdpop",       # one weight per age stratum
  standstrata = "age_group",    # single standardisation variable
  by          = "group",        # single grouping variable
  ref         = "A",            # reference group by label
  mult        = 100000
)
```

---

## Case 2 — Standardise by age AND sex jointly

Pass a character vector to `standstrata` to standardise by more than one
variable simultaneously. Here we use age (3 bands) × sex (2 levels) = **6
joint strata**.

**Key requirement:** `stdpop` must contain the *joint* standard-population
weight for each (age × sex) combination — not separate marginal weights.
The weights within each group must sum to a positive constant (they are
normalised internally).

```{r}
age_sex_data <- data.frame(
  age_group = rep(rep(c("18-39","40-64","65+"), each=2), 2),
  sex       = rep(c("F","M"), 6),
  ethnicity = rep(c("Asian","White"), each=6),
  events    = c( 4, 5, 12, 15,  9, 12,
                10,12, 28, 35, 22, 30),
  population= c(1900,2000,1700,1600,320,300,
                4800,5000,4200,4000,900,800),
  stdpop    = rep(c(0.20,0.20, 0.20,0.20, 0.10,0.10), 2)
)

directsrr(
  data        = age_sex_data,
  event       = "events",
  pop         = "population",
  stdpop      = "stdpop",
  standstrata = c("age_group", "sex"),   # joint standardisation
  by          = "ethnicity",
  ref         = "White",
  mult        = 100000
)
```

---

## Case 3 — Stratify by two grouping variables

Pass a character vector to `by` to produce one rate per combination of
grouping variables. When `by` has multiple columns, `ref` must be a named
list specifying a value for **every** `by` column.

```{r}
two_group_data <- data.frame(
  age_group = rep(rep(c("18-39","40-64","65+"), each=2), 4),
  sex       = rep(c("F","M"), 12),
  ethnicity = rep(c("Asian","Asian","White","White"), each=6),
  region    = rep(c("North","South","North","South"), each=6),
  events    = c( 2, 3,  7, 9,  9,12,
                 5, 6, 14,18, 16,22,
                 7, 8, 18,20, 22,30,
                12,15, 32,40, 40,55),
  population= c(1400,1500,1300,1200, 220,200,
                1900,2000,1900,1800, 380,350,
                2900,3000,2600,2500, 550,500,
                3800,4000,3700,3500, 800,700),
  stdpop    = rep(c(0.20,0.20, 0.20,0.20, 0.10,0.10), 4)
)

directsrr(
  data        = two_group_data,
  event       = "events",
  pop         = "population",
  stdpop      = "stdpop",
  standstrata = c("age_group", "sex"),
  by          = c("ethnicity", "region"),
  ref         = list(ethnicity = "White", region = "North"),  # named list
  mult        = 100000,
  digits_rate  = 1,
  digits_ratio = 3
)
```

---

## Specifying the reference group: `ref` vs `ref_index`

Two separate arguments remove any ambiguity about what the reference is.

Use **`ref`** to identify by label — the most common approach:

```{r eval=FALSE}
# Character label
directsrr(..., ref = "Control")

# Numeric label — matches the group whose by-column value equals 2,
# not the second row
directsrr(..., ref = 2)

# Named list for multi-column by
directsrr(..., ref = list(arm = "Placebo", centre = "London"))
```

Use **`ref_index`** to identify by row position in the output table —
no integer-literal tricks required:

```{r eval=FALSE}
directsrr(..., ref_index = 1)   # first row  (the default)
directsrr(..., ref_index = 2)   # second row
```

Supplying both `ref` and `ref_index` is an error. When neither is given
and `by` is not `NULL`, the first output row is used as reference.

---

## Output columns

| Column | Description |
|--------|-------------|
| `crude` | Crude (unstandardised) rate × `mult` |
| `rateadj` | Directly standardised rate × `mult` |
| `lb_gam` / `ub_gam` | Tiwari et al. (2006) gamma CI |
| `se_gam` | SE from the Tiwari variance |
| `lb_dob` / `ub_dob` | Dobson et al. (1991) CI (`ub_dob = NA` for zero-event groups; use `ub_gam` instead) |
| `srr` | Standardised rate ratio vs reference (`NA` if reference rate is zero) |
| `lb_srr` / `ub_srr` | F-distribution CI for the SRR |
| `srd` | Standardised rate difference vs reference (always computed) |
| `lb_srd` / `ub_srd` | Normal CI for the SRD |

## Controlling zero-event notifications

When `dobson = TRUE` and a group has zero events, `ub_dob` is `NA`.
The `notify` argument controls how this is signalled:

```{r eval=FALSE}
directsrr(..., notify = "warn")     # default — warning(), catchable with tryCatch()
directsrr(..., notify = "message")  # message(), suppressible with suppressMessages()
directsrr(..., notify = "silent")   # no output — useful in loops or simulations
```

## Merging the standard population

```{r eval=FALSE}
std <- data.frame(
  age_band   = c("18-39", "40-64", "65+"),
  esp_weight = c(0.4, 0.4, 0.2)
)
mydata <- merge(mydata, std, by = "age_band")
```

For joint age-sex standardisation, derive joint weights from marginals:

```{r eval=FALSE}
age_w <- c("18-39"=0.40, "40-64"=0.40, "65+"=0.20)
sex_w <- c("F"=0.50, "M"=0.50)

std_joint <- expand.grid(age_band=names(age_w), sex=names(sex_w))
std_joint$stdpop <- age_w[std_joint$age_band] * sex_w[std_joint$sex]
mydata <- merge(mydata, std_joint, by = c("age_band","sex"))
```

## Single pooled rate

Omit `by` for one overall rate with no comparison:

```{r}
directsrr(
  data        = directsrr_ExampleData,
  event       = "events",
  pop         = "population",
  stdpop      = "stdpop",
  standstrata = "age_group",
  mult        = 100000
)
```

## References

Tiwari RC, Clegg LX, Zou Z (2006). Efficient interval estimation for age-adjusted cancer rates. *Statistical Methods in Medical Research*, **15**(6), 547-569.

Fay MP, Feuer EJ (1997). Confidence intervals for directly standardized rates: A method based on the gamma distribution. *Statistics in Medicine*, **16**(7), 791-801. DOI: 10.1002/(SICI)1097-0258(19970415)16:7<791::AID-SIM500>3.0.CO;2-#

Dobson AJ, Kuulasmaa K, Eberle E, Scherer J (1991). Confidence intervals for weighted sums of Poisson parameters. *Statistics in Medicine*, **10**(3), 457-462. DOI: 10.1002/sim.4780100317

Consonni D, Coviello E, Buzzoni C, Mensi C (2012). A command to calculate age-standardized rates with efficient interval estimation. *Stata Journal*, **12**(4), 688-701.

---

## Individual participant data (IPD)

`directsrr()` accepts both **aggregate** and **individual-level** data.
The internal `aggregate(..., FUN = sum)` step collapses rows by
(group x age stratum) before any calculation, so IPD is handled automatically.

For IPD set `event` to a 0/1 column and `pop` to each participant's
person-time (e.g. years of follow-up). Merge the standard-population weight
onto the IPD by age stratum first.

```{r eval=FALSE}
# IPD: one row per participant
std_pop <- data.frame(age_group = c("18-39","40-64","65+"),
                      stdpop    = c(0.4, 0.4, 0.2))
ipd <- merge(my_individual_data, std_pop, by = "age_group")
# columns needed: age_group, event (0/1), follow_up_years, stdpop, group

directsrr(ipd,
         event       = "event",
         pop         = "follow_up_years",
         stdpop      = "stdpop",
         standstrata = "age_group",
         by          = "group",
         ref         = "Control",
         mult        = 100000)
```

**Float person-time** is handled correctly throughout. The `N` column in
the output preserves decimal places (e.g. 12.8 person-years) rather than
silently rounding to 13. Use `digits_count` to control display precision.

**Survey-weighted data:** if event counts are fractional after applying
weights (e.g. 2.7 events), the Tiwari gamma CI (`lb_gam` / `ub_gam`)
remains valid. The Dobson CI assumes integer counts; a warning is issued if
non-integer totals are detected — rely on `ub_gam` in that case.
